---
title: "gvmap tutorial"
author: "Dai Yuting"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{gvmap}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

## DESCRIPTION

Gvmap is an R package that can integrate multiple heatmap and legend figures. 

## INSTALLATION

Gvmap depends on several R packages. They are 
[configr](https://cran.r-project.org/web/packages/configr/index.html), 
[dendextend](https://cran.r-project.org/web/packages/dendextend/index.html), 
[matrixStats](https://cran.r-project.org/web/packages/matrixStats/index.html), 
[easySVG](https://github.com/ytdai/easySVG), 
[rsvg](https://cran.r-project.org/web/packages/rsvg/index.html), and 
[stringr](https://cran.r-project.org/web/packages/stringr/index.html). 

You can install gvmap through github or source code.


### Github

``` {r echo = TRUE, eval = FALSE}
# Install the cutting edge development version from GitHub:
# install.packages("devtools")

# install easySVG
devtools::install_github("ytdai/easySVG")

# install other packages
install.packages("configr")
install.packages("dendextend")
install.packages("matrixStats")
install.packages("rsvg")
install.packages("stringr")

devtools::install_github("ytdai/gvmap")
```

### Source code

``` {r echo = TRUE, eval = FALSE}
install.packages("gvmap_0.1.0.tar.gz")
```

## Prepare heatmap data

The heatmap data must be a list or a matrix, and the name of the list must like this form: heatmap_1, heatmap_2 and so on.

The colnames of each single heatmap data must be the sample, and the rownames can be genes.

**ATTENTION**: heatmap_1 is the main heatmap. This setting is default and there is **ONLY ONE** main heatmap in this figure.

``` {r}
heatmap_file <- system.file("extdata/heatmap.txt", package = "gvmap")

heatmap_data_1 <- read.table(heatmap_file, header = T)
rownames(heatmap_data_1) <- heatmap_data_1$X
heatmap_data_1 <- heatmap_data_1[, -1]

head(heatmap_data_1)

heatmap_data_2 <- heatmap_data_1[20:50, ]
heatmap_data_3 <- heatmap_data_1[80:100, ]

# if you have multiple heatmaps, the heatmap data must be a list
# heatmap_1 is the main heatmap
heatmap_data_mtp <- list(heatmap_1 = heatmap_data_1,
                     heatmap_2 = heatmap_data_2,
                     heatmap_3 = heatmap_data_3)

# if you only have one heatmap, the heatmap data can be a matrix or a list
# matrix
heatmap_data_sig <- heatmap_data_1
# list
heatmap_data_sig <- list(heatmap_1 = heatmap_data_1)

```

## Prepare legend data

The legend is the supplement, annotation and explanation of heatmap samples. The rownames of legend data 
must be the sample, and the colnames must be the attributes.

``` {r}
legend_file <- system.file("extdata/legend.txt", package = "gvmap")

legend_data <- read.table(legend_file, header = T)
row.names(legend_data) <- legend_data$SampleName

legend_data
```

## Prepare config file

The config file must be a [YAML](https://en.wikipedia.org/wiki/YAML) format. SEE SECTION **Config paramters**

``` {r}
config_file <- system.file("extdata/config.yaml", package = "gvmap")

```

## Run gvmap

``` {r}
library(configr)
library(dendextend)
library(matrixStats)
library(easySVG)
library(rsvg)
library(stringr)
library(gvmap)

# run by default paramter
config_file <- system.file("extdata/config.mtp.yaml", package = "gvmap")

gvmap(legend_data = legend_data,
      heatmap_data = heatmap_data_mtp,
      config_file = config_file,
      output_svg_name = "o1.svg",
      plot_width = 800,
      plot_height = 1200)
```

![](o1.svg)

``` {r}
# run by default paramter
config_file <- system.file("extdata/config.sig.yaml", package = "gvmap")

gvmap(legend_data = legend_data,
      heatmap_data = heatmap_data_sig,
      config_file = config_file,
      output_svg_name = "o2.svg",
      plot_width = 800,
      plot_height = 1200)
```

![](o2.svg)

``` {r}
# only legend data
config_file <- system.file("extdata/config.leg.yaml", package = "gvmap")

gvmap(legend_data = legend_data,
      config_file = config_file,
      output_svg_name = "o3.svg",
      plot_width = 800,
      plot_height = 1200)
```

![](o3.svg)
``` {r}
# add sample span and heatmap span
config_file <- system.file("extdata/config.mtp.yaml", package = "gvmap")

gvmap(legend_data = legend_data,
      heatmap_data = heatmap_data_mtp,
      config_file = config_file,
      output_svg_name = "o4.svg",
      sample_span = 5,
      heatmap_row_span = 4,
      plot_width = 800,
      plot_height = 1200)
```


![](o4.svg)

## Config paramters

``` {r echo = TRUE, eval = FALSE}
# the config file need to be like this
map_config:
  heatmap_num: 2
  legend_num: 2
  heatmap_1:
    raw_data: FALSE
    kmer_col: 3
    kmer_row: 4
    split_row_name: NULL
    percentage: 0.4
    kmer_col_color: NULL
    kmer_row_color: "pool_col"
    dendrogram: "both"
    Rowv: TRUE
    Colv: TRUE
    color_theme: "gradient_col_gr"
    distfun: "dist"
    hclustfun: "hclust"
  heatmap_2:
    raw_data: TRUE
    split_row_name: NULL
    percentage: 0.1
    dendrogram: "both"
    Rowv: TRUE
    color_theme: "gradient_col_op"
    distfun: "dist"
    hclustfun: "hclust" 
  heatmap_3:
    percentage: 0.1
  legend_1:
    - ["Gender","binary_col", 3]
    - ["Age", "binary_col", 2]
    - ["Fusion", "tag_col", 4]
    - ["other_fusion", "pool_col", 5]
  legend_2:
    - ["KRAS", "mutation_col", 6]
    - ["P53", "mutation_col", 7]
    - ["NRAS", "mutation_col", 8]
  map_order:
    - "heatmap_1"
    - "legend_1"
    - "heatmap_2"
    - "heatmap_3"
    - "legend_2"
  split_sample: 
    - "M1003"
    - "M4002"

color_config:
  bg_col: "#EEEEEE"
  tag_col: "#696969"
  binary_col: ["#196ABD", "#C20B01"]
  white_col: "#FFFFFF"
  black_col: "#000000"
  mutation_col:
    missense: "#3987CC"
    frameshift: "#DB3D3D"
    nonsense: "#FF7F0E"
    protein_del: "#7F7F7F"
    protein_ins: "#8C564B"
    one_hit: "#819981"
    tow_hit: "#499E49"
    three_hit: "#237023"
  pool_col: ["#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D", "#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A"]
  gradient_col_gr: ["green", "black", "red"]
  gradient_col_yp: ["yellow", "black", "purple"]
  gradient_col_op: ["orange", "white", "purple"]
  gv_line_col: "#FDBF6F"
  gv_bar_col: "#E6AB02"
  gv_dot_col: "#000000"
  kmer_col: ["green", "blue", "red"]

```

And basic parameters: 

### PARAMETERS IN MAP_CONFIG

**======= HEATMAP ========**

[param] `raw_data`  TRUE or FALSE, determines whether to change the raw data to row z-data to
illustrate the heatmap.

[param] `kmer_col`  number, split the col data by kmer. ATTENTION: only heatmap_1 has this parameter

[param] `kmer_row`  number, split the row data by kmer.

[param] `split_row_name`  vector, the specific row name to split, eg. ["name1", "name2", "name3"]

[param] `percentage`  number, between 0 and 1, the total heatmap percentage of the whole canvas.

[param] `kmer_col_color`  character. Col data color theme of dendrogram.

[param] `kmer_row_color`  character. Row data color them of dendrogram.

[param] `dendrogram`  "both", "col", "row", "none".

[param] `color_theme`  character heatmap gradient colot theme.

[param] `Rowv`  TRUE or FALSE, determines if and how the row dendrogram should be reordered.	By default,
it is TRUE, which implies dendrogram is computed and reordered based on row means.
If NULL or FALSE, then no dendrogram is computed and no reordering is done. If a
dendrogram, then it is used "as-is", ie without any reordering. If a vector of integers,
then dendrogram is computed and reordered based on the order of the vector.

[param] `Colv`  TRUE or FALSE, determines if and how the column dendrogram should be reordered.	Has the options
as the Rowv argument above and additionally when x is a square matrix, Colv = "Rowv" means
that columns should be treated identically to the rows. 

[param] `distfun`  character, function used to compute the distance (dissimilarity) between both rows
and columns. Defaults to dist. 

[param] `hclustfun`  character, function used to compute the hierarchical clustering when Rowv or
Colv are not dendrograms. Defaults to hclust. SEE \link[gplots]{heatmap.2}

**======= LEGEND ========**

the legend data config format is simple

eg:
- ["Gender","binary_gender_col", 3]

- ["ETV6-RUNX1", "tag_col", 4]

- ["KRAS", "mutation_col", 7]

The first element in the vector is legend data. It dose not the same as the colnames in legend data.

The second element in the vector is the color theme, you can choose the color from color config list.

The third element in the vector is the column number.

**======= OTHER ========**

[param] `heatmap_num` the total heatmap number

[param] `legend_num` the total legend number

[param] `map_order` the order of all plot element

[param] `split_sample` split sample name


